---
title: "CodeChal7"
author: "Jaden King"
output: 
  md_document: 
  word_document: 
    toc: true
date: "2025-04-03"
---

```{r setup, include=FALSE}

```

1.	4 pts. Read in the data called “PlantEmergence.csv” using a relative file path and load the following libraries. tidyverse, lme4, emmeans, multcomp, and multcompView. Turn the Treatment , DaysAfterPlanting  and Rep into factors using the function as.factor
```{r}
#load required packages
library(tidyverse)
library(lme4)
library(emmeans)
library(multcomp)
library(multcompView)

#read in the data
plantemerg <- read.csv("PlantEmergence.csv")
head(plantemerg)

plantemerg$Treatment <- as.factor(plantemerg$Treatment)
plantemerg$DaysAfterPlanting <- as.factor(plantemerg$DaysAfterPlanting)
plantemerg$Rep <- as.factor(plantemerg$Rep)

str(plantemerg) #ensure transformations worked
```

 
2.	5 pts. Fit a linear model to predict Emergence using Treatment and DaysAfterPlanting along with the interaction. Provide the summary of the linear model and ANOVA results. 
```{r}
lm1 <- lm(Emergence~Treatment*DaysAfterPlanting, data= plantemerg)
```


3.	5 pts. Based on the results of the linear model in question 2, do you need to fit the interaction term? Provide a simplified linear model without the interaction term but still testing both main effects. Provide the summary and ANOVA results. Then, interpret the intercept and the coefficient for Treatment 2.

with interaction
```{r}
summary(lm1)

```

- The interaction between days after planting and rep was not significant and does not need to be included in the model. 

without interation
```{r}
lm_model <- lm(Emergence~Treatment+DaysAfterPlanting, data= plantemerg)
summary(lm_model)
anova(lm_model)
```

Interpretation of Results- both significant
- Intercept coefficient = 182.163 
  - predicted value when x = 0
- Treatment 2 coefficient = -134.531
  - Trt 2 is 134.531 [y units] less emergence than trt 1


4.	5 pts. Calculate the least square means for Treatment using the emmeans package and perform a Tukey separation with the compact letter display using the cld function. Interpret the results.
```{r}
lsmeans <- emmeans(lm_model, ~ Treatment)
results_lsmeans <- cld(lsmeans, alpha=0.5, reversed= TRUE, details= TRUE)
results_lsmeans
```
Results
- Treatment 5 had the highest emmergence
- Treatment 2 had the lowest emergence
- Treatment 5,3,6,8 are not statistically different from each other
- Treatment 3,6,8,4 are not statistically different from each other
- Treatment 5 is statistically higher than 4,9,1,7,2
- Treatment 3 is statistically higher than 9,1,7,2
- Treatment 6 and 8 are statistically higher than 1,7,2
- All treatments are statistically higher than trt 2


5.	4 pts. The provided function lets you dynamically add a linear model plus one factor from that model and plots a bar chart with letters denoting treatment differences. Use this model to generate the plot shown below. Explain the significance of the letters. 
```{r}
plot_cldbars_onefactor <- function(lm_model, factor) {
  data <- lm_model$model
  variables <- colnames(lm_model$model)
  dependent_var <- variables[1]
  independent_var <- variables[2:length(variables)]

  lsmeans <- emmeans(lm_model, as.formula(paste("~", factor))) # estimate lsmeans 
  Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE, Letters = letters) # contrast with Tukey adjustment by default.
  
  # Extracting the letters for the bars
  sig.diff.letters <- data.frame(Results_lsmeans$emmeans[,1], 
                                 str_trim(Results_lsmeans$emmeans[,7]))
  colnames(sig.diff.letters) <- c(factor, "Letters")
  
  # for plotting with letters from significance test
  ave_stand2 <- lm_model$model %>%
    group_by(!!sym(factor)) %>%
    dplyr::summarize(
      ave.emerge = mean(.data[[dependent_var]], na.rm = TRUE),
      se = sd(.data[[dependent_var]]) / sqrt(n())
    ) %>%
    left_join(sig.diff.letters, by = factor) %>%
    mutate(letter_position = ave.emerge + 10 * se)
  
  plot <- ggplot(data, aes(x = !! sym(factor), y = !! sym(dependent_var))) + 
    stat_summary(fun = mean, geom = "bar") +
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +
    ylab("Number of emerged plants") + 
    geom_jitter(width = 0.02, alpha = 0.5) +
    geom_text(data = ave_stand2, aes(label = Letters, y = letter_position), size = 5) +
    xlab(as.character(factor)) +
    theme_classic()
  
  return(plot)
}


plot_cldbars_onefactor(lm_model, "Treatment")
```

The plot shows the number of emerged plants on y axis and treatment on x axis. The letters above the bars represent the statistical grouping. Treatments with the same letter are not significantly different from each other. By having the letters on the plot, it is easier to see differences between treatments. 