---
title: "JadenKing_DataWrangle_CodeChal"
output:
  word_document: default
  html_document: default
---
requires tidyverse package
```{r}
library(tidyverse)
```

read data in
```{r}
metadata <- read.csv("metadata.csv")
diversity <- read.csv("DiversityData.csv")
```

connect the data frames by column code
```{r}
alpha <- data.frame(left_join(metadata, diversity, by = "Code"))

head(alpha)
```

Calculate Pielou’s evenness index: Pielou’s evenness is an ecological parameter calculated by the Shannon diversity index (column Shannon) divided by the log of the richness column. 
-	Using mutate, create a new column to calculate Pielou’s evenness index. 
-	Name the resulting dataframe alpha_even.

```{r}
alpha_even <- data.frame(mutate(alpha, Peven = shannon/(log(richness))))

head(alpha_even)
```

Using tidyverse language of functions and the pipe, use the summarise function and tell me the mean and standard error evenness grouped by crop over time.
-	Start with the alpha_even dataframe
-	Group the data: group the data by Crop and Time_Point.
-	Summarize the data: Calculate the mean, count, standard deviation, and standard error for the even variable within each group.
-	Name the resulting dataframe alpha_average
```{r}
alpha_average<- data.frame(alpha_even %>%
  group_by(Crop, Time_Point) %>%
  summarize(Mean.even = mean(Peven),
            n=n(),
            sd.dev = sd(Peven))%>%
  mutate(std.err = sd.dev/sqrt(n)))

head(alpha_average)
```

Calculate the difference between the soybean column, the soil column, and the difference between the cotton column and the soil column
-	Start with the alpha_average dataframe
-	Select relevant columns: select the columns Time_Point, Crop, and mean.even.
-	Reshape the data: Use the pivot_wider function to transform the data from long to wide format, creating new columns for each Crop with values from mean.even.
-	Calculate differences: Create new columns named diff.cotton.even and diff.soybean.even by calculating the difference between Soil and Cotton, and Soil and Soybean, respectively.
-	Name the resulting dataframe alpha_average2
```{r}
alpha_average2 <- data.frame(alpha_average %>%
  select(Time_Point,Crop,Mean.even) %>%
  pivot_wider(names_from = Crop, values_from = Mean.even) %>%
  mutate(diff.cotton.even = Soil - Cotton) %>%
  mutate(diff.soybean.even = Soil - Soybean))

head(alpha_average2)
```

Connecting it to plots
- Start with the alpha_average2 dataframe
- Select relevant columns: select the columns Time_Point, diff.cotton.even, and diff.soybean.even.
-	Reshape the data: Use the pivot_longer function to transform the data from wide to long format, creating a new column named diff that contains the values from diff.cotton.even and diff.soybean.even
-	Create the plot: Use ggplot and geom_line() with ‘Time_Point’ on the x-axis, the column ‘values’ on the y-axis, and different colors for each ‘diff’ category. The column named ‘values’ come from the pivot_longer. The resulting plot should look like the one to the right. 
```{r}
alpha_average2 %>%
  select(Time_Point, diff.cotton.even, diff.soybean.even) %>%
  pivot_longer(c(diff.cotton.even, diff.soybean.even),names_to="diff") %>%
  ggplot(aes(x= Time_Point, y= value, color= diff)) +
  geom_line() +
  theme_classic() +
  xlab("Time (hrs)") +
  ylab("Difference from soil in Pielou's evenness")
```

